---
title: "Reproducible Research: Peer Assessment 1"
author: "Alex Sickert"
date: "July 18, 2015"
output: html_document
---


## Loading and preprocessing the data


```{r, echo=TRUE}
myData <-read.csv("activity.csv", sep = ",", header = TRUE, as.is = TRUE)
```


## What is mean total number of steps taken per day?

```{r, echo=TRUE}
stepsPerDay <- aggregate(myData["steps"], by=myData[c("date")], FUN=sum)
# print(stepsPerDay)
```



```{r, echo=TRUE}
res <- aggregate(myData["steps"], by=myData[c("date")], FUN=sum)
myHist =  res[,2]
hist(myHist, main="Histogram", xlab="Number of steps per day", ylab="Frequency")
```


###Mean and Median calculation

Median:

```{r, echo=TRUE}
myMedian = median(myHist, na.rm = TRUE)
print("Median:")
print(myMedian)

```

Mean:

```{r, echo=TRUE}

myMean = mean(myHist, na.rm = TRUE)
print("Mean:")
print(myMean)
```

## What is the average daily activity pattern?

```{r, echo=TRUE}
res <- aggregate(myData["steps"], by=myData[c("interval")], FUN=mean, na.rm=TRUE)
# print(res)
plot(res$interval, res$steps, type="l", xlab= "Interval", ylab= "Mean steps",  lwd=2)
```

Interval with most steps
```{r, echo=TRUE}
rowIndex = which.max( res[,2] )
print(res[rowIndex,])
```

## Imputing missing values

Number of missing steps

```{r, echo=TRUE}
missing = sum(is.na(myData$steps))
print(missing)
```


Strategey to fill values: If there is a NA then we use the value of the interval before. And if the interval before is also NA then we use zero

```{r, echo=TRUE}
myFilledData = myData
lastValue = 0 
for(i in 1:length(myFilledData$steps)){
  if(is.na(myFilledData[i,1]) ){
    myFilledData[i,1] = lastValue
  }else{
    lastValue = myFilledData[i,1]
  }
}
```


Historgram for filled dataset 

```{r, echo=TRUE}
resFilled <- aggregate(myFilledData["steps"], by=myFilledData[c("date")], FUN=sum)
myHistFilled =  resFilled[,2]
hist(myHistFilled, main="Histogram with filled data", xlab="Number of steps per day", ylab="Frequency")
```



Mean and Median calculation for filled dataset

```{r, echo=TRUE}
myMedianFilled = median(myHistFilled, na.rm = TRUE)
print("Median:")
print(myMedianFilled)
myMeanFilled = mean(myHistFilled, na.rm = TRUE)
print("Mean:")
print(myMeanFilled)
```


## Are there differences in activity patterns between weekdays and weekends?

```{r, echo=TRUE}
library(timeDate)
library(lattice)
library(datasets)

myFilledData$weekend <- factor(isWeekend(as.Date(myFilledData$date)))
myFilledData$weekend <- factor(myFilledData$weekend, labels = c("Weekday", "Weekend"))

res <- aggregate(myFilledData["steps"], by=myFilledData[c("interval", "weekend")], FUN=mean, na.rm=TRUE)

xyplot(steps ~ interval | weekend , data = res, type = "l", layout = c(1,2))
```

